---
title: "MOMI Data Analysis"
output:  
  html_document:
    toc: true               # Enable table of contents
    toc_float: true         # Floating TOC (sticks to side)
    toc_depth: 3            # Show up to 3 heading levels
    code_folding: hide      # Hide code by default
    theme: cosmo            # Optional: adds styling
    highlight: tango 
---

# MOMI Data Analysis

## EDA

### Loading Required Libraries

```{r chunk_1, message=FALSE, warning=FALSE}
# Load required packages
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
#library(skimr)
library(patchwork)
library(gtsummary)
```

### Importing MOMI Data

#### 1. Sample Maternal Data

```{r chunk_2, message=FALSE, warning=FALSE}
sample_maternal_data <- read_excel("C:/Users/Dell/Desktop/THSTI-CDSA work/MOMI Data Analysis/preg_intensities_sb_stratified_maternal.xlsx", col_types = "text")

sample_maternal_data <- sample_maternal_data %>%
    type.convert(as.is = TRUE)  # Automatically convert to appropriate types

head(sample_maternal_data)    
```

```{r chunk_3, message=FALSE, warning=FALSE}
cat("Dimesions of Maternal Data:",dim(sample_maternal_data))
```

#### 2. Sample cord data

```{r chunk_4, message=FALSE, warning=FALSE}
sample_cord_data <- read_excel("C:/Users/Dell/Desktop/THSTI-CDSA work/MOMI Data Analysis/preg_intensities_sb_stratified_cord.xlsx", col_types = "text")

# Convert columns to appropriate types
sample_cord_data <- sample_cord_data %>%
    type.convert(as.is = TRUE)

head(sample_cord_data)
```

```{r chunk_5, message=FALSE, warning=FALSE}
cat("Dimesions of cord Data:",dim(sample_cord_data))
```

#### 3. Clinical Data

```{r chunk_6, message=FALSE, warning=FALSE}
clinical_data <- read_excel("C:/Users/Dell/Desktop/THSTI-CDSA work/MOMI Data Analysis/momi_combined_data.xlsx", col_types = "text")

clinical_data <- clinical_data %>%
    type.convert(as.is = TRUE)  # Automatically convert to appropriate types

head(clinical_data)
```

```{r chunk_7, message=FALSE, warning=FALSE}
cat("Dimesions of clinical data:",dim(clinical_data))
```

### Adding Site and GA_weeks in clinical data

#### Adding Site from Site Code

```{r chunk_8, message=FALSE, warning=FALSE}
clinical_data <- clinical_data %>%
  mutate(SITE = case_when(
    SITE_CODE == 1 ~ "ZAPPS",
    SITE_CODE == 2 ~ "GAPPSB",
    SITE_CODE == 3 ~ "AMANHIP",
    SITE_CODE == 4 ~ "AMANHIB",
    SITE_CODE == 5 ~ "AMANHIT",
    SITE_CODE == 6 ~ "THSTI",
    TRUE ~ as.character(SITE_CODE)  # Handle any other cases
  ))

  head(clinical_data$SITE)
```

```{r chunk_9, message=FALSE, warning=FALSE}
# Check for missing values in SITE column
sum(is.na(clinical_data$SITE))
```

#### Calculating GA_weeks to merge both files

```{r chunk_10, message=FALSE, warning=FALSE}
# Add new columns to the clinical_data dataframe
clinical_data <- clinical_data %>%
  mutate(
    VISITDT = as.Date(VISITDT, format = "%Y-%m-%d", origin = "1899-12-30"),
    VDT_ENR = as.Date(VDT_ENR, format = "%Y-%m-%d", origin = "1899-12-30"),
    DATE_LMP = as.Date(DATE_LMP, format = "%Y-%m-%d", origin = "1899-12-30"),
    DATE_USG = as.Date(DATE_USG, format = "%Y-%m-%d", origin = "1899-12-30"),
    
    # Create a unique ID by combining ORIG_ID and SITE
    id = paste(ORIG_ID, SITE, sep = "_"),
    
    # Calculate GA_crl (Gestational Age based on Crown-Rump Length)
    GA_crl = exp(1.684969 + (0.315646 * CRL) - (0.049306 * CRL^2) + 
                 (0.004057 * CRL^3) - (0.000120456 * CRL^4)),
    
    # Calculate GA_hadlock (Gestational Age based on Hadlock formula)
    GA_hadlock = 10.50 + 0.197 * (BPD * FL) + 0.9500 * FL + 0.7300 * BPD,
    
    # Calculate GA_weeks (final gestational age at enrollment)
    GA_weeks = ifelse(is.na(GA_crl), GA_hadlock, GA_crl)
  )

head(clinical_data$GA_weeks)
```

#### Summary of GA_weeks and its derivatives

```{r chunk_11, message=FALSE, warning=FALSE}
clinical_data %>%
  select(CRL, FL, BPD, GA_crl, GA_hadlock, GA_weeks) %>%
  tbl_summary(
    statistic = list(all_continuous() ~ "Median : {median}, IQR: ({p25}, {p75}), Values present: {N_nonmiss}, NA : {N_miss}"),
    missing = "no"
  ) %>%
  as_kable()
```

## Data cleaning

### Formatting Columns

```{r chunk_12, message=FALSE, warning=FALSE}
colnames(clinical_data)
colnames(sample_maternal_data)
colnames(sample_cord_data)
```

```{r chunk_13, message=FALSE, warning=FALSE}
# Ensure column names match
colnames(sample_maternal_data) <- toupper(colnames(sample_maternal_data))
colnames(clinical_data) <- toupper(colnames(clinical_data))
colnames(sample_cord_data) <- toupper(colnames(sample_cord_data))

# common columns in both maternal and clinical data
common_columns <- intersect(colnames(sample_maternal_data), colnames(clinical_data))
print(common_columns)
```

### Duplicate rows

```{r chunk_14, message=FALSE, warning=FALSE}
# Find duplicate rows based on ORIG_ID, SITE, VISITDT and BABY_ID
duplicate_rows <- clinical_data %>%
  group_by(ORIG_ID, SITE, VISITDT, BABY_ID) %>% 
  filter(n() > 1) %>% # Keep only rows that appear more than once
  ungroup() 

print(duplicate_rows)
cat("Total duplicate rows:", nrow(duplicate_rows), "\n")
```

## Data Distribution

### Calculating Participants

```{r chunk_15, message=FALSE, warning=FALSE}
#Unique count of distinct columns (ORIG_ID, SITE) in both data frames
clinical_particip <- clinical_data %>%
  distinct(ORIG_ID, SITE) %>%
  nrow()

maternal_particip <- sample_maternal_data %>%
  distinct(ORIG_ID, SITE) %>%
  nrow()

cord_particip <- sample_cord_data %>%
  distinct(ORIG_ID, SITE) %>%
  nrow()

cat("Participants in clinical data:",clinical_particip,"\n")
cat("Participants in maternal data:",maternal_particip,"\n")
cat("Participants in cord data:",cord_particip,"\n")
```

#### Count of Maternal participants`s clinical data available or not

```{r chunk_16, message=FALSE, warning=FALSE}
matching_rows <- suppressMessages(inner_join(
    sample_maternal_data %>% distinct(ORIG_ID, SITE),
    clinical_data %>% distinct(ORIG_ID, SITE),
    relationship = "many-to-many"
) %>% nrow())

# Find records in maternal data that don't match clinical data
maternal_no_match <- suppressMessages(anti_join(
    sample_maternal_data %>% distinct(ORIG_ID, SITE),
    clinical_data %>% distinct(ORIG_ID, SITE)
) %>% nrow())

cat("Participants (clinical data available):",matching_rows,"\n")
cat("Participants those clinical data is not present:", maternal_no_match)
```

#### Count of cord participants`s clinical data available or not

```{r chunk_17, message=FALSE, warning=FALSE}
matching_rows <- suppressMessages(inner_join(
    sample_cord_data %>% distinct(ORIG_ID, SITE),
    clinical_data %>% distinct(ORIG_ID, SITE),
    relationship = "many-to-many"
) %>% nrow())

# Find records in cord data that don't match clinical data
cord_no_match <- suppressMessages(anti_join(
    sample_cord_data %>% distinct(ORIG_ID, SITE),
    clinical_data %>% distinct(ORIG_ID, SITE)
) %>% nrow())

cat("Participants (clinical data available):", matching_rows, "\n")
cat("Participants whose clinical data is not present:", cord_no_match, "\n")
```

#### Participants matching in maternal and cord data

```{r chunk_18, message=FALSE, warning=FALSE}
matching_rows <- suppressMessages(inner_join(
    sample_cord_data %>% distinct(ORIG_ID, SITE),
    sample_maternal_data %>% distinct(ORIG_ID, SITE),
    relationship = "many-to-many"
) %>% nrow())

# Find records in cord data that don't match maternal data
maternal_no_match <- suppressMessages(anti_join(
    sample_cord_data %>% distinct(ORIG_ID, SITE),
    sample_maternal_data %>% distinct(ORIG_ID, SITE)
) %>% nrow())

cat("Participants available in both data:",matching_rows,"\n")
cat("Participants (from cord data) not available in maternal data:", maternal_no_match)
```

#### Formatting in Table

```{r chunk_19, message=FALSE, warning=FALSE}
# Create a data frame
data_table <- data.frame(
  Data = c("Clinical data", "Maternal data", "Cord data"),
  Participants = c(clinical_particip, maternal_particip, cord_particip),
  Rows = c(nrow(clinical_data), nrow(sample_maternal_data), nrow(sample_cord_data)),
  Columns = c(ncol(clinical_data), ncol(sample_maternal_data), ncol(sample_cord_data)),
  stringsAsFactors = FALSE # Avoid automatic factor conversion
)

# Convert the data frame to a tibble
data_table_tibble <- as_tibble(data_table)

# Print the tibble
print(data_table_tibble)
```

### Discrepancy between clinical and metabolomics datasets

#### Maternal sample data

```{r chunk_20, message=FALSE, warning=FALSE}
# Participants in maternal data
total_maternal <- sample_maternal_data %>% distinct(ORIG_ID, SITE) %>% nrow()

# Find records in maternal sample data that don't match clinical data
overall_maternal_no_match <- suppressMessages(anti_join(
    sample_maternal_data %>% distinct(ORIG_ID, SITE),
    clinical_data %>% distinct(ORIG_ID, SITE)
) %>% nrow())

cat("Maternal sample data:", "\n")
cat("Total records in sample_maternal_data:", total_maternal, "\n")
cat("Records not present in clinical_data (Sample only):", overall_maternal_no_match, "\n")
```

#### Clinical data

```{r chunk_21, message=FALSE, warning=FALSE}
# Participants in clinical data
total_clinical <- clinical_data %>% distinct(ORIG_ID, SITE) %>% nrow()

# Find records in clinical data that don't match maternal sample data
overall_clinical_no_match <- suppressMessages(anti_join(
    clinical_data %>% distinct(ORIG_ID, SITE),
    sample_maternal_data %>% distinct(ORIG_ID, SITE)
) %>% nrow())

cat("Clinical data:", "\n")
cat("Total records in clinical data:", total_clinical, "\n")
cat("Records not present in sample data (Clinical only):", overall_clinical_no_match, "\n")
```

#### Sample and clinical

```{r chunk_22, message=FALSE, warning=FALSE}
# Find matching records between sample_maternal_data and clinical_data
matching_data <- inner_join(
    sample_maternal_data,
    clinical_data,
    by = c("SITE", "ORIG_ID"),
    relationship = "many-to-many"
)  %>%
distinct(SITE, ORIG_ID)

sample_clinical <- nrow(matching_data)

cat("Matching Data Analysis:\n")
cat("Total matching records:", sample_clinical, "\n")
```

#### Total 

```{r chunk_23, message=FALSE, warning=FALSE}
# full outer join to get total data in Sample and Clinical
total_data <- full_join(
    sample_maternal_data,
    clinical_data,
    by = c("SITE", "ORIG_ID"),
    relationship = "many-to-many"
) %>%
distinct(SITE, ORIG_ID)

# Calculate total matches
total_matches <- nrow(total_data)

# Display matching analysis
cat("Matching Data Analysis:\n")
cat("Total matching records (distinct SITE and ORIG_ID):", total_matches, "\n")
```

#### Discrepancy BY SITES

```{r chunk_24, message=FALSE, warning=FALSE}
# Get unique SITE values
sites <- unique(c(sample_maternal_data$SITE, clinical_data$SITE))

sample_only <- c()
clinical_only <- c()
sample_and_clinical <- c()
total <- c()

# Create analysis for each site
for(site in sites) {
    # Filter data for current site
    maternal_site <- sample_maternal_data %>% 
        filter(SITE == site) %>% 
        distinct(ORIG_ID, SITE)
    
    clinical_site <- clinical_data %>% 
        filter(SITE == site) %>% 
        distinct(ORIG_ID, SITE)
    
    # Calculate statistics
    total_maternal <- nrow(maternal_site)
    total_clinical <- nrow(clinical_site)
    
    # Find unmatched records (suppress join messages)
    maternal_no_match <- suppressMessages(
        anti_join(maternal_site, clinical_site) %>% nrow()
    )
    clinical_no_match <- suppressMessages(
        anti_join(clinical_site, maternal_site) %>% nrow()
    )
    
    # Find matching records (suppress join message)
    matching <- suppressMessages(
        inner_join(maternal_site, clinical_site) %>% nrow()
    )

    # Total distinct records in both datasets
    total_matching <- suppressMessages(
        full_join(maternal_site, clinical_site) %>% nrow())
    
    # Display results for current site
    cat("\nAnalysis for", site, ":\n")
      # Append site only if it is not already in sites
    if (!(site %in% sites)) {
        sites <- c(sites, site)
    }
    cat("Maternal Data:\n")
    cat("Total distinct records:", total_maternal, "\n")
    cat("Records not in clinical data (Sample only):", maternal_no_match, "\n")
    sample_only <- c(sample_only, maternal_no_match)
    
    cat("Clinical Data:\n")
    cat("Total distinct records:", total_clinical, "\n")
    cat("Records not in maternal data (Clinical only):", clinical_no_match, "\n")
    clinical_only <- c(clinical_only, clinical_no_match)
    
    cat("Matching Records (Sample and Clinical):", matching, "\n")
    sample_and_clinical <- c(sample_and_clinical, matching)
    cat("TOtal distinct records in both Sample and Clinical:", total_matching, "\n")
    total <- c(total, total_matching)
    cat("----------------------------------------\n")
}
```

#### Formatting in Table

```{r chunk_25, message=FALSE, warning=FALSE}
# Create a data frame using the vectors
data_table <- data.frame(
  sites = c(sites, "OVERALL"),
  sample_only = c(sample_only, overall_maternal_no_match),
  clinical_only = c(clinical_only, overall_clinical_no_match),
  sample_and_clinical = c(sample_and_clinical, sample_clinical),
  total = c(total, total_matches)
)

# Convert the data frame to a tibble
data_table_tibble <- as_tibble(data_table)

# Print the tibble
print(data_table_tibble)
```

### Visit wise clinical data distribution

```{r chunk_26, message=FALSE, warning=FALSE}
# Find duplicate rows based on ORIG_ID, SITE, and VISITDT
duplicate_rows <- clinical_data %>%
  group_by(ORIG_ID, SITE, VISITDT, BABY_ID) %>% # Group by the unique key columns
  filter(n() > 1) %>% # Keep only rows that appear more than once
  ungroup() %>% # Remove grouping
  select(ORIG_ID, SITE, VISITDT, BABY_ID) # Select only the relevant columns

# Print the duplicate rows
print(duplicate_rows)
```

#### Adding Visit column in Clinical Data

```{r chunk_27, message=FALSE, warning=FALSE}
# Add a visit_number column to clinical_data
clinical_data <- clinical_data %>%
  select(ORIG_ID, SITE, VISITDT, BABY_ID, everything()) %>% # Ensure VISITDT is included
  group_by(ORIG_ID, SITE, BABY_ID) %>% # Group by unique key columns
  arrange(VISITDT, .by_group = TRUE) %>% # Arrange VISITDT in ascending order for each group
  mutate(visit_number = row_number()) %>% # Assign visit numbers based on the order of VISITDT
  ungroup() # Remove grouping
```

```{r chunk_28, message=FALSE, warning=FALSE}
# Initialize an empty list to store visit statistics
visit <- list()

# Loop through visits 1 to 4
for (i in 1:4) {
  # Filter data for the current visit
  visits <- clinical_data %>% filter(visit_number == i)
  
  # Print the number of participants for the current visit
  cat("Visit", i, ":", nrow(visits), "\n")
  
  # If there are no rows for the current visit, skip the statistics calculation
  if (nrow(visits) == 0) {
    visit[[i]] <- list(
      participants = 0,
      min_max = "No data",
      mean_sd = "No data",
      median_iqr = "No data"
    )
    next
  }
  
  # Calculate statistics for GA_WEEKS
  min_val <- min(visits$GA_WEEKS, na.rm = TRUE)
  max_val <- max(visits$GA_WEEKS, na.rm = TRUE)
  mean_val <- mean(visits$GA_WEEKS, na.rm = TRUE)
  sd_val <- sd(visits$GA_WEEKS, na.rm = TRUE)
  median_val <- median(visits$GA_WEEKS, na.rm = TRUE)
  q1 <- quantile(visits$GA_WEEKS, 0.25, na.rm = TRUE) # 25th percentile
  q3 <- quantile(visits$GA_WEEKS, 0.75, na.rm = TRUE) # 75th percentile
  
  # Store statistics in the desired format
  visit[[i]] <- list(
    participants = nrow(visits),
    min_max = paste(round(min_val, 0), round(max_val, 0), sep = ", "),
    mean_sd = paste(round(mean_val, 0), "(", round(sd_val, 0), ")", sep = ""),
    median_iqr = paste(round(median_val, 0), "(", round(q1, 0), "-", round(q3, 0), ")", sep = "")
  )
}
```

#### Formatting in Table

```{r chunk_29, message=FALSE, warning=FALSE}
# Initialize the data frame with the first column
data_table <- data.frame(
  total = c("Participants (N)", "GA (WEEKS) MIN, MAX", "GA WEEKS Mean(SD)", "GA WEEKS Median (IQR)")
)

# Loop through the visits and add columns dynamically
for (i in 1:4) {
  if (!is.null(visit[[i]])) {
    data_table[[paste0("visit", i)]] <- c(
      visit[[i]]$participants, # Number of participants
      visit[[i]]$min_max, # Min, Max
      visit[[i]]$mean_sd, # Mean(SD)
      visit[[i]]$median_iqr # Median(IQR)
    )
  } else {
    data_table[[paste0("visit", i)]] <- c("No data", "No data", "No data", "No data")
  }
}

# Convert the data frame to a tibble
data_table_tibble <- as_tibble(data_table)

# Print the tibble
print(data_table_tibble)
```

### Distribution of GA by Visit

```{r chunk_30, message=FALSE, warning=FALSE}
plots <- list()

for (i in 1:4) {
  visit <- clinical_data %>% filter(visit_number == i, !is.na(GA_WEEKS))
  
  p <- ggplot(data = visit, aes(x = GA_WEEKS)) +
    geom_histogram(binwidth = 1, fill = "blue", color = "black") +
    labs(
      title = paste("Distribution of GA by Visit", i),
      x = "Gestational Age WEEKS",
      y = "Frequency"
    ) +
    theme_minimal()
  
  # Add the plot to the list
  plots[[i]] <- p
}

# Combine all plots using patchwork
combined_plot <- plots[[1]] + plots[[2]] + plots[[3]] + plots[[4]] + 
  plot_layout(ncol = 2) # Arrange plots in 2 columns

print(combined_plot)
```

### Trimester wise maternal sample distribution

#### Adding Trimester Column

```{r chunk_31, message=FALSE, warning=FALSE}
# Add a trimester column to clinical_data
clinical_data <- clinical_data %>%
  mutate(
    TRIMESTER = case_when(
      is.na(GA_WEEKS) ~ NA_integer_,  # Handle missing values
      GA_WEEKS < 14 ~ 1,  # First trimester
      GA_WEEKS < 28 ~ 2,  # Second trimester
      TRUE ~ 3            # Third trimester
    )
  )

# Display the first few rows to verify
head(clinical_data$TRIMESTER)
```

```{r chunk_32, message=FALSE, warning=FALSE}
clinical_data %>%
  select(GA_WEEKS, TRIMESTER) %>%
  head()
```

## Removing NA values

```{r chunk_33, message=FALSE, warning=FALSE}
columns <- c("ORIG_ID", "SITE", "GA_WEEKS")

for(col in columns) {
    cat("NA count in", col, ":", sum(is.na(clinical_data[[col]])), "\n")
}
```

```{r chunk_34, message=FALSE, warning=FALSE}
# Store original dimensions
orig_dim <- dim(clinical_data)

# Filter out rows with missing values in key columns
clinical_data <- clinical_data %>%
  filter(!is.na(GA_WEEKS))

# Display filtering results
cat("Original dimensions:", orig_dim[1], "rows,", orig_dim[2], "columns\n")
cat("After filtering:", nrow(clinical_data), "rows,", ncol(clinical_data), "columns\n")
cat("Rows removed:", orig_dim[1] - nrow(clinical_data), "\n")

# Verify no missing values in key columns
cat("\nMissing values:\n")
cat("GA_WEEKS:", sum(is.na(clinical_data$GA_WEEKS)), "\n")
```

```{r chunk_35, message=FALSE, warning=FALSE}
columns <- c("ORIG_ID", "SITE", "GA_WEEKS")

for(col in columns) {
    cat("NA count in", col, ":", sum(is.na(clinical_data[[col]])), "\n")
}
```

## Merging clinical and sample data

#### Changing Data Types of both Dataframes

```{r chunk_36, message=FALSE, warning=FALSE}
# Check data types
cat("Sample data:\n")
cat("ORIG_ID types:", class(sample_maternal_data$ORIG_ID), "\n")
cat("SITE types:", class(sample_maternal_data$SITE), "\n")
cat("GA_WEEKS types:", class(sample_maternal_data$GA_WEEKS), "\n")

cat("\nClinical data:\n")
cat("ORIG_ID types:", class(clinical_data$ORIG_ID), "\n")
cat("SITE types:", class(clinical_data$SITE), "\n")
cat("GA_WEEKS types:", class(clinical_data$GA_WEEKS), "\n")
```

```{r chunk_37, message=FALSE, warning=FALSE}
# Convert columns to same type in both dataframes
sample_maternal_data <- sample_maternal_data %>%
  mutate(
    ORIG_ID = as.character(ORIG_ID),
    SITE = as.character(SITE),
    GA_WEEKS = as.numeric(GA_WEEKS)
  )

clinical_data <- clinical_data %>%
  mutate(
    ORIG_ID = as.character(ORIG_ID),
    SITE = as.character(SITE),
    GA_WEEKS = as.numeric(GA_WEEKS)
  )

  # Check data types
cat("Sample data:\n")
cat("ORIG_ID types:", class(sample_maternal_data$ORIG_ID), "\n")
cat("SITE types:", class(sample_maternal_data$SITE), "\n")
cat("GA_WEEKS types:", class(sample_maternal_data$GA_WEEKS), "\n")

cat("\nClinical data:\n")
cat("ORIG_ID types:", class(clinical_data$ORIG_ID), "\n")
cat("SITE types:", class(clinical_data$SITE), "\n")
cat("GA_WEEKS types:", class(clinical_data$GA_WEEKS), "\n")
```

#### Removing Unwanted spaces

```{r chunk_38, message=FALSE, warning=FALSE}
# Check for any case differences or leading/trailing spaces
clinical_data <- clinical_data %>%
  mutate(
    ORIG_ID = trimws(ORIG_ID),
    SITE = trimws(SITE),
    GA_WEEKS = round(GA_WEEKS, 2)  # Round to 2 decimal places
  )

sample_maternal_data <- sample_maternal_data %>%
  mutate(
    ORIG_ID = trimws(ORIG_ID),
    SITE = trimws(SITE),
    GA_WEEKS = round(GA_WEEKS, 2)  # Round to 2 decimal places
  )
```

#### Left Join

```{r chunk_39, message=FALSE, warning=FALSE}
dim(clinical_data)
```

```{r chunk_40, message=FALSE, warning=FALSE}
# Perform left join
leftjoin_data <- merge(
    x = clinical_data,
    y = sample_maternal_data,
    by = c("ORIG_ID", "SITE", "GA_WEEKS"),
    all.x = TRUE    # Keep all rows from clinical_data
)

# Display dimensions and summary
cat("Dimensions of datasets:\n")
cat("Clinical data:", dim(clinical_data), "\n")
cat("Sample maternal data:", dim(sample_maternal_data), "\n")
cat("Merged data:", dim(leftjoin_data), "\n")
```

```{r chunk_41, message=FALSE, warning=FALSE}
cat("NA values in SB_CASE in maternal data:", sum(is.na(sample_maternal_data$SB_CASE)), "\n")
```

```{r chunk_42, message=FALSE, warning=FALSE}
cat("Total rows:", nrow(leftjoin_data), "\n")
cat("NA values in merged data in SB_CASE:", sum(is.na(leftjoin_data$SB_CASE)), "\n")
cat("Non-NA values in merged data in SB_CASE:", nrow(leftjoin_data) - sum(is.na(leftjoin_data$SB_CASE)), "\n")
```

#### Inner Join

```{r chunk_43, message=FALSE, warning=FALSE}
# Perform inner join to get only matching records
matched_data <- inner_join(clinical_data,sample_maternal_data,
                by = c("ORIG_ID", "SITE", "GA_WEEKS"))

# Display dimensions and summary
cat("Dimensions of datasets:\n")
cat("Clinical data:", dim(clinical_data), "\n")
cat("Sample maternal data:", dim(sample_maternal_data), "\n")
cat("Matched data:", dim(matched_data), "\n")

head(matched_data)
```

```{r chunk_44, message=FALSE, warning=FALSE}

```

